---
title: "p8105_hw6_jdv2118"
author: "Justin Vargas"
output: github_document
---

# Loading Libraries and Setting Graphic Settings

```{r}
library(tidyverse)

library(viridis)

library(modelr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d

scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

To obtain a distribution for $\hat{r}^2$, we'll follow basically the same procedure we used for regression coefficients: draw bootstrap samples; the a model to each; extract the value I'm concerned with; and summarize. Here, we'll use `modelr::bootstrap` to draw the samples and `broom::glance` to produce `r.squared` values. 

```{r weather_df, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


```{r}
weather_df %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  ggplot(aes(x = r.squared)) + geom_density()
```

In this example, the $\hat{r}^2$ value is high, and the upper bound at 1 may be a cause for the generally skewed shape of the distribution. If we wanted to construct a confidence interval for $R^2$, we could take the 2.5% and 97.5% quantiles of the estimates across bootstrap samples. However, because the shape isn't symmetric, using the mean +/- 1.96 times the standard error probably wouldn't work well.

We can produce a distribution for $\log(\beta_0 * \beta1)$ using a similar approach, with a bit more wrangling before we make our plot.

```{r}
weather_df %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  select(id = `.id`, term, estimate) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate) %>% 
  rename(beta0 = `(Intercept)`, beta1 = tmin) %>% 
  mutate(log_b0b1 = log(beta0 * beta1)) %>% 
  ggplot(aes(x = log_b0b1)) + geom_density()
```

As with $r^2$, this distribution is somewhat skewed and has some outliers. 

The point of this is not to say you should always use the bootstrap -- it's possible to establish "large sample" distributions for strange parameters / values / summaries in a lot of cases, and those are great to have. But it is helpful to know that there's a way to do inference even in tough cases. 

# Problem 2

```{r}
homicide_data = 
  read_csv("data/homicide-data.csv") %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    result = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Closed by arrest"      ~ 1,
      disposition == "Open/No arrest"        ~ 0
      ),
    victim_age = as.numeric(victim_age)
  ) %>% 
  filter(city_state != "Dallas, TX",
         city_state != "Phoenix, AZ",
         city_state != "Kansas City, MO",
         city_state != "Tulsa, AL",
victim_race %in% c("White", "Black"),
victim_sex != "Unknown") %>% 
  select(city_state, result, victim_age, victim_race, victim_sex)
```

```{r}
baltimore_glm =
  homicide_data %>% 
  filter(city_state == "Baltimore, MD")

glm(result ~ victim_age + victim_sex + victim_race, 
    data = baltimore_glm,
    family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate),
    ci_lower = exp(estimate - 1.96 * std.error),
    ci_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, OR, starts_with("ci")) %>% 
  knitr::kable(digits = 3)
```

```{r}
cities_glm = 
  homicide_data %>% 
  nest(data = -city_state) %>% 
  mutate(
    city_models = 
      map(.x = data, ~glm(result ~ victim_age + victim_sex + victim_race, data = .x, family = binomial())),
    result = map(city_models, broom::tidy)
  ) %>% 
  select(city_state, result) %>% 
  unnest(result) %>% 
  mutate(
    OR = exp(estimate),
    ci_lower = exp(estimate - 1.96 * std.error),
    ci_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("ci")) %>%
  filter(term == "victim_sexMale")
  
cities_glm
```

```{r}
cities_plot =
  cities_glm %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper)) + 
  labs(
    x = "City and State Names",
    y = "Odds Ratio",
    title = "The Odds Ratio for Each City and State"
  ) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

cities_plot
```

# Problem 3

```{r}
birthweight_data =
  read_csv("data/birthweight.csv") %>%
  mutate(
    babysex = as.factor(babysex),
    babysex = recode(babysex, "1" = "Male", "2" = "Female"),
    frace = as.factor(frace),
    frace = recode(frace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other",      "9" = "Unknown"),
    malform = as.factor(malform),
    malform = recode(malform, "0" = "Absent", "1" = "Present"),
    mrace = as.factor(mrace),
    mrace = recode(mrace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other")
  )

head(birthweight_data)

sum(is.na(birthweight_data))
```

```{r}
birthweight_full_model = lm(bwt ~ ., data = birthweight_data)

birthweight_full_model
```

```{r}
birthweight_plot =
birthweight_data %>% 
  modelr::add_predictions(., birthweight_full_model) %>% 
  modelr::add_residuals(., birthweight_full_model) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .2) +
  labs(
    x = "Fitted Values",
    y = "Residuals",
    title = "Residuals vs Fitted Values"
  )

birthweight_plot
```

